### 2021-07-03 ch2flipFlop.tex
made some progress on the nightmare citations.

A relativity weak flow shearing a bed of sediment entrains individual particles into a state of motion controlled by turbulent forcing and intermittent collisions with other grains at rest on the bed, generating wide fluctuations in the sediment velocity \citep{Heyman2016,Fathel2015}.
Bed load particles move downstream until they are disentrained when they happen to encounter sufficiently sheltered pockets on the bed surface to interrupt their motions \citep{Charru2004,Gordon1972}.
Eventually, the bed around them rearranges and destroys this shelter, or turbulent fluctuations overcome the shelter \citep{Celik2019,Valyrakis2014}, particles are once again entrained, and the cycle repeats.
Bed load transport is thus a kind of itinerant motion, characterized by alternation between fluctuating movements and rest.
This process is difficult to describe mathematically given the technicality of the stochastic physics involved \citep{Furbish2017,Ancey2020}.

@@

A set of studies has developed stochastic formulations of the sediment flux which improve on the mean field description provided by Einstein \citep{Furbish2012}. 
Sediment fluxes generally exhibit wide fluctuations due to variations in the number of moving particles and their velocities \citep{Ancey2008,Ancey2014}, the migration of bedforms and sediment waves \citep{Hamamori1962,Guala2014,Hoey1992,Recking2012}, and a host of other processes \citep{Dhont2018}.
Because these fluctuations occur over disparate timescales, measurements of mean sediment fluxes depend on the timescale over which they are collected, a phenomenon called scale-dependence \citep{Saletti2014,Dhont2019,Singh2009,Turowski2010,Ancey2020}.
To date, very few models have calculated the probability distribution of the bed load sediment flux due to individual particle motions \citep{Ancey2008,Ancey2014}, and among these, even fewer have described any observation-scale dependence of the flux \citep{Ancey2020,Turowski2010}. Among those that do, none have yet formulated the sediment flux in terms of the downstream trajectories of individual grains.

@@

\section{Results \label{sec:res}}
\subsection{The position probability distribution and its moments}
\begin{figure}
	\centerline{\includegraphics{figures/ch2/figure2_slopeKey.pdf}}
	\caption{Panel (a) indicates the probability distribution of particle position as it evolves through time. From the initial mixture of motion and rest states, particles advect downstream as they diffuse due to differences in their fluctuating velocities and exchange between motion and rest. Panel (b) shows the resulting particle diffusion. At timescales $t \ll 2D/V^2$, the diffusion is normal since the movement is approximately a standard diffusion process (as advection is irrelevant on these timescales). For larger timescales, $2D/V^2 \ll t \ll \tau$, particles undergo ballistic diffusion similar to \citet{Lisle1998} as a result of some particles being stationary as others advect. Finally longer than the timescale $\tau = 1/k$ associated with entrainment and deposition, diffusion is again normal, exemplified by exchange between motion and rest. All results are scaled by the mean hop length $l=V/k_D$ and the autocorrelation time $\tau=1/k$ of the motion/rest alternation.}
	\label{fig:fig1}
\end{figure}
The master equation \ref{eq:master} describes the evolution of the probability distribution of position through time. Loosely speaking the solution of this equation is expected to be some combination of the \citet{Einstein1937} theory for particle transport with the Gaussian propagator of the advection-diffusion equation \citep{Gillespie1992}.

@@

This chapter has generalized the earlier descriptions of individual sediment trajectories \citep[e.g.][]{Lisle1998,Lajeunesse2018} to include velocity fluctuations in the motion state. Using results from this generalized model as an example, I demonstrated how to calculate the sediment flux probability distribution, phrasing earlier renewal theory approaches more directly in terms of the underlying particle dynamics \citep[e.g.][]{Turowski2010,Ancey2020}. 
Calculated in this way, the sediment flux distribution is observation-scale dependent, describing how sediment fluxes depend on the period over which they are measured.

@@

The sediment flux probability distribution in equation \ref{eq:fluxdist} represents a Poisson distribution with an observation-scale dependent rate.
Poisson distributions have a relatively thin tail, meaning fluctuations are typically small \citep{Ancey2006}.
In reality, sediment flux distributions are only Poissonian at high transport rates, whereas in other conditions they have wide tails representing the possibility of extremely large transport fluctuations \citep{Ancey2008,Turowski2010,Dhont2010,Saletti2015} which appear as bursts \citep[e.g.][]{Goh2008} in the sediment flux timeseries \citep{Singh2009, Heyman2013,Benavides2021}. This exposes a need to generalize a mechanistic theory of the sediment flux I developed here to produce wider fluctuations.

Descriptions of sediment transport based on population dynamics in a control volume have realistically-wide fluctuations by incorporating a positive feedback between the number of moving particles and the particle entrainment rate called collective entrainment \citep{Ancey2008,Ancey2014}.
This feedback generates spatially-correlated waves of moving particles \citep{Ancey2014, Heyman2015} and produces non-exponential interarrival time distributions \citep{Heyman2013} which imply wide-tailed flux distributions when incorporated in renewal theory \citep{Turowski2010,Ancey2020}.
Collective entrainment has been attributed to particle-particle interactions, such as small granular avalanches and collision-induced entrainment, and to fluid-particle interactions, such as large eddies entraining particles as they sweep downstream \citep{Heyman2014b,Ancey2014,Lee2018}.

We can imagine generalizing the description in equation \ref{eq:Langevin} to include such fluid-particle interactions by including time dependence in the movement velocity ($V$), diffusivity ($D$), or entrainment and deposition rates ($k_E$ and $k_D$) as necessary, although the resulting equations would likely require numerical solutions.
Particle-particle interactions would be more challenging to include.
The dynamical equation \ref{eq:langevin} would generalize to $\dot{x}_i(t) = F_i(x_i,t) + \sum_{i\neq j}G_{ij}(\{x_j\},t)$, where the $F_i$ are the driving terms unique to each particle, while the $G_{ij}$ are some (generally stochastic) terms representing interactions between the $i$th and $j$th particles \citep{Goldstein1956}. The resulting joint distribution of particle positions and velocities ($P(x_1,v_1,\dots,x_{N(t)},v_{N(t)},t)$) might be formulated by analogy to the theory of reaction diffusion systems \citep{Cardy2006}, non-ideal gases \citep{Chapman1970,Brilliantov2004}, or other interacting particle models available in physics literature \citep{Hernandez2004,Escaff2018}. 
By analogy to collective entrainment, interacting models which generalize the approach established in this chapter should be capable of producing wide sediment transport fluctuations.

@@

Considering the flux in the Exner equation as a stochastic quantity requires us to reinterpret channel dynamics in a way which few researchers have explored \citep{Jerolmack2005,Bohorquez2014}.
In contrast to the traditional analysis, a stochastic Exner equation with fluctuating $q(x,t)$ generates an ensemble of possibilities for the channel profile $z(x)$ at each instant. This set of possibilities is characterized by a probability distribution $P(\{z(x)\},t)$ for each possible profile.
Mathematically, since the channel profile is a field, evaluating such a probability distribution of channel profiles would require statistical field theory \citep{Kardar2007}.
A starting point on this analysis might be $P(\{z(x)\},t) = \bra \delta[z(x,t)-h(x,t)]\ket,$ where $h(x,t)$ involves the stochastic flux $q(x,t)$ from equation \ref{eq:stocEx}.
In this equation, $\delta[f(x)]$ is a Dirac delta functional, which generalizes the delta function $\delta(x)$ from a single-valued parameter $x$ to a continuous field $f(x)$.
Such ideas are commonly applied to understand polymers and membranes \citep{Kawakatsu2004,Nelson2004}, but they have not been introduced in geomorphology.

Introducing such a description with the stochastic Exner equation would lend new mathematical language to the old unsolved problem in geomorphology: how does variability affect landscape evolution?
The classic works all acknowledge variability \citep{Leopold1952}, although one of their main efforts was to construct strategies to describe landscapes without explicitly accounting for it, producing ideas like competent flows to summarize climatic variability \citep{Wolman1960}, characteristic grain sizes to overcome sorting processes \citep{Parker1982}, or average flow energies to avoid turbulence \citep{Bagnold1954}.
Progress in stochastic approaches to river science \citep[e.g.][]{Furbish2021c,Ancey2020b} invites us to step beyond averaged descriptions, propagate noises through the governing equations of landscape evolution, and ask again with these classic works to what extent fluctuations shape Earth's surface.

@@

These results generalize the bedload trajectory models of \citet{Einstein1937}, \citet{Lisle1998}, and \citet{Lajeunesse2018} to include fluctuating movement velocities, and they reframe earlier renewal models of the sediment flux probability distribution \citep{Lajeunesse1010,Ancey2020} to explicitly include the transport dynamics of individual sediment grains. They further quantify the dependence of the sediment flux on the observation scale, and provide guidance as to the measurement times required to resolve sediment transport without ambiguity, at least in simplified configurations when bedforms are absent. 
Next steps are to include interactions between particles to produce wider sediment transport fluctuations, and to evaluate the effects of sediment transport fluctuations on channel morphodynamics.

>>>>

A relativity weak flow shearing a bed of sediment entrains individual particles into a state of motion controlled by turbulent forcing and intermittent collisions with other grains at rest on the bed, generating wide fluctuations in the sediment velocity \citep{Heyman2016,Fathel2015}.
Bed load particles move downstream until they are disentrained when they happen to encounter sufficiently sheltered pockets on the bed surface to interrupt their motions \citep{Charru2004,Gordon1972}.
Eventually, the bed around them rearranges and destroys this shelter, or turbulent fluctuations overcome the shelter \citep{Celik2014,Valyrakis2010}, particles are once again entrained, and the cycle repeats.
Bed load transport is thus a kind of itinerant motion, characterized by alternation between fluctuating movements and rest.
This process is difficult to describe mathematically given the technicality of the stochastic physics involved \citep{Furbish2017,Ancey2020}.

@@

A set of studies has developed stochastic formulations of the sediment flux which improve on the mean field description provided by Einstein \citep{Furbish2012}. 
Sediment fluxes generally exhibit wide fluctuations due to variations in the number of moving particles and their velocities \citep{Ancey2008,Ancey2014}, the migration of bedforms and sediment waves \citep{Hamamori1962,Guala2014,Hoey1992,Recking2012}, and a host of other processes \citep{Dhont2018}.
Because these fluctuations occur over disparate timescales, measurements of mean sediment fluxes depend on the timescale over which they are collected, a phenomenon called scale-dependence \citep{Saletti2015,Dhont2018,Singh2009,Turowski2010,Ancey2020}.
To date, very few models have calculated the probability distribution of the bed load sediment flux due to individual particle motions \citep{Ancey2008,Ancey2014}, and among these, even fewer have described any observation-scale dependence of the flux \citep{Ancey2020,Turowski2010}. Among those that do, none have yet formulated the sediment flux in terms of the downstream trajectories of individual grains.

@@

\section{Results \label{sec:res}}
\subsection{The position probability distribution and its moments}
\begin{figure}
	\centerline{\includegraphics{figures/ch2/figure2_slopeKey.pdf}}
	\caption{Panel (a) indicates the probability distribution of particle position as it evolves through time. From the initial mixture of motion and rest states, particles advect downstream as they diffuse due to differences in their fluctuating velocities and exchange between motion and rest. Panel (b) shows the resulting particle diffusion. At timescales $t \ll 2D/V^2$, the diffusion is normal since the movement is approximately a standard diffusion process (as advection is irrelevant on these timescales). For larger timescales, $2D/V^2 \ll t \ll \tau$, particles undergo ballistic diffusion similar to \citet{Lisle1998} as a result of some particles being stationary as others advect. Finally longer than the timescale $\tau = 1/k$ associated with entrainment and deposition, diffusion is again normal, exemplified by exchange between motion and rest. All results are scaled by the mean hop length $l=V/k_D$ and the autocorrelation time $\tau=1/k$ of the motion/rest alternation.}
	\label{fig:fig1}
\end{figure}
The master equation \ref{eq:master} describes the evolution of the probability distribution of position through time. Loosely speaking the solution of this equation is expected to be some combination of the \citet{Einstein1937} theory for particle transport with the Gaussian propagator of the advection-diffusion equation \citep{Morse1953a}.

@@

This chapter has generalized the earlier descriptions of individual sediment trajectories \citep[e.g.][]{Lisle1998,Lajeunesse2017} to include velocity fluctuations in the motion state. Using results from this generalized model as an example, I demonstrated how to calculate the sediment flux probability distribution, phrasing earlier renewal theory approaches more directly in terms of the underlying particle dynamics \citep[e.g.][]{Turowski2010,Ancey2020}. 
Calculated in this way, the sediment flux distribution is observation-scale dependent, describing how sediment fluxes depend on the period over which they are measured.

@@

The sediment flux probability distribution in equation \ref{eq:fluxdist} represents a Poisson distribution with an observation-scale dependent rate.
Poisson distributions have a relatively thin tail, meaning fluctuations are typically small \citep{Ancey2006}.
In reality, sediment flux distributions are only Poissonian at high transport rates, whereas in other conditions they have wide tails representing the possibility of extremely large transport fluctuations \citep{Ancey2008,Turowski2010,Dhont2018,Saletti2015} which appear as bursts \citep[e.g.][]{Goh2008} in the sediment flux timeseries \citep{Singh2009, Heyman2013,Benavides2021}. This exposes a need to generalize a mechanistic theory of the sediment flux I developed here to produce wider fluctuations.

Descriptions of sediment transport based on population dynamics in a control volume have realistically-wide fluctuations by incorporating a positive feedback between the number of moving particles and the particle entrainment rate called collective entrainment \citep{Ancey2008,Ancey2014}.
This feedback generates spatially-correlated waves of moving particles \citep{Ancey2014, Heyman2015} and produces non-exponential interarrival time distributions \citep{Heyman2013} which imply wide-tailed flux distributions when incorporated in renewal theory \citep{Turowski2010,Ancey2020}.
Collective entrainment has been attributed to particle-particle interactions, such as small granular avalanches and collision-induced entrainment, and to fluid-particle interactions, such as large eddies entraining particles as they sweep downstream \citep{Heyman2014a,Ancey2014,Lee2018}.

We can imagine generalizing the description in equation \ref{eq:Langevin} to include such fluid-particle interactions by including time dependence in the movement velocity ($V$), diffusivity ($D$), or entrainment and deposition rates ($k_E$ and $k_D$) as necessary, although the resulting equations would likely require numerical solutions.
Particle-particle interactions would be more challenging to include.
The dynamical equation \ref{eq:langevin} would generalize to $\dot{x}_i(t) = F_i(x_i,t) + \sum_{i\neq j}G_{ij}(\{x_j\},t)$, where the $F_i$ are the driving terms unique to each particle, while the $G_{ij}$ are some (generally stochastic) terms representing interactions between the $i$th and $j$th particles \citep{Goldstein1997}. The resulting joint distribution of particle positions and velocities ($P(x_1,v_1,\dots,x_{N(t)},v_{N(t)},t)$) might be formulated by analogy to the theory of reaction diffusion systems \citep{Pechenik1999, Cardy2008}, non-ideal gases \citep{Chapman1970,Brilliantov2004}, or other interacting particle models available in physics literature \citep{Hernandez2004,Escaff2018}. 
By analogy to collective entrainment, interacting models which generalize the approach established in this chapter should be capable of producing wide sediment transport fluctuations.

@@

Considering the flux in the Exner equation as a stochastic quantity requires us to reinterpret channel dynamics in a way which few researchers have explored \citep{Jerolmack2005,Bohorquez2016}.
In contrast to the traditional analysis, a stochastic Exner equation with fluctuating $q(x,t)$ generates an ensemble of possibilities for the channel profile $z(x)$ at each instant. This set of possibilities is characterized by a probability distribution $P(\{z(x)\},t)$ for each possible profile.
Mathematically, since the channel profile is a field, evaluating such a probability distribution of channel profiles would require statistical field theory \citep{Kardar2007}.
A starting point on this analysis might be $P(\{z(x)\},t) = \bra \delta[z(x,t)-h(x,t)]\ket,$ where $h(x,t)$ involves the stochastic flux $q(x,t)$ from equation \ref{eq:stocEx}.
In this equation, $\delta[f(x)]$ is a Dirac delta functional, which generalizes the delta function $\delta(x)$ from a single-valued parameter $x$ to a continuous field $f(x)$.
Such ideas are commonly applied to understand polymers and membranes \citep{Kawakatsu2001,Nelson2004}, but they have not been introduced in geomorphology.

Introducing such a description with the stochastic Exner equation would lend new mathematical language to the old unsolved problem in geomorphology: how does variability affect landscape evolution?
The classic works all acknowledge variability \citep{Leopold1952}, although one of their main efforts was to construct strategies to describe landscapes without explicitly accounting for it, producing ideas like competent flows to summarize climatic variability \citep{Wolman1959,Wolman1978}, characteristic grain sizes to overcome sorting processes \citep{Parker1982}, or average flow energies to avoid turbulence \citep{Bagnold1954}.
Progress in stochastic approaches to river science \citep[e.g.][]{Furbish2021c,Ancey2020b} invites us to step beyond averaged descriptions, propagate noises through the governing equations of landscape evolution, and ask again with these classic works to what extent fluctuations shape Earth's surface.

@@

These results generalize the bedload trajectory models of \citet{Einstein1937}, \citet{Lisle1998}, and \citet{Lajeunesse2017} to include fluctuating movement velocities, and they reframe earlier renewal models of the sediment flux probability distribution \citep{Lajeunesse2010,Ancey2020} to explicitly include the transport dynamics of individual sediment grains. They further quantify the dependence of the sediment flux on the observation scale, and provide guidance as to the measurement times required to resolve sediment transport without ambiguity, at least in simplified configurations when bedforms are absent. 
Next steps are to include interactions between particles to produce wider sediment transport fluctuations, and to evaluate the effects of sediment transport fluctuations on channel morphodynamics.

### 2021-07-03 ch3bedEle.tex
made some progress on the nightmare citations.

Resting time distributions have been carefully studied in relation to these models because the predicted diffusion characteristics are critically dependent on whether the distribution has a light or heavy tail \citep{Bradley2017,Martin2012,Weeks1998}.
Resting times have puzzled researchers because early experiments show exponential distributions \citep{Einstein1937,Hubbell1964, Yano1969,Nakagawa1976}, while later experiments show heavy-tailed power-law distributions \citep{Martin2012,Voepel2013, Olinde2015, Pretzlav2016, Bradley2017, Liu2019}.
A predominant hypothesis is that power-law distributed resting times originate from buried grains \citep{Martin2014,Voepel2013}; this hypothesis permits surface grains to retain exponential resting times. 
Conceptually, when grains rest on the surface, material transported from upstream can deposit on top of them, preventing entrainment until its removal, driving up resting times and imparting a heavy tail to the distribution.
To our knowledge, \citet{Martin2014} have provided the only direct support for this hypothesis by tracking grains through complete cycles of burial and exhumation using a narrow flume with glass walls.
They observed heavy-tailed resting times of buried grains and described their results with a mathematical model similar to an earlier effort by \citet{Voepel2013}.
Both of these models treat bed elevation changes as a random walk and interpret resting times as return periods from above in the bed elevation time-series \citep{Redner2007}. Each describes resting time distributions from different experiments, but they rely on different random walk models, and their treatment of bed elevations as a process independent of sediment transport is questionable at first glance, since bedload transport is the source of bed elevation changes \citep{Wong2007}, and neither model explicitly includes bedload transport.
Models of sedimentary bed evolution incorporating sediment transport processes might enhance understanding of sediment resting times.

The second type of stochastic model prescribes rates (probabilities per unit time) to the erosion and deposition events of individual grains within a control volume to calculate the particle activity \citep{Einstein1950}.
These approaches aim at a complete statistical characterization of the bedload flux \citep{Furbish2012a,Fathel2015,Furbish2017,Heyman2016}, including probability distributions \citep{Ancey2006,Ancey2008}, spatial and temporal characteristics of its fluctuations \citep{Ancey2008, Heyman2014a, Roseberry2012,Dhont2018}, and the dependence of these statistical characteristics on the length and time scales over which they are measured or calculated \citep{Singh2009,Singh2012,Saletti2015a,Ma2014}.
A recent surge in research activity has generated rapid progress and spawned many new inquiries in this subject. 
For example, \citet{Ancey2006} demonstrated that a constant erosion rate as originally proposed by \citet{Einstein1950} was insufficient to develop realistically large particle activity fluctuations, so they added a positive feedback between the particle activity and erosion rate they called ``collective entrainment'' \citep{Ancey2008, Heyman2013,Heyman2014,Ma2014, Lee2018}.
While they deemed this feedback necessary to model realistic activity fluctuations, the implications of this collective entrainment term on bed topography and particle activity changes has not been fully explored.

@@

Bed elevation changes modify the likelihood of entrainment and deposition in a negative feedback \citep{Sawai1987, Wong2007}; that is, aggradation increases the likelihood of entrainment, while degradation increases the likelihood of deposition.
\citet{Wong2007} concluded that bed elevation changes induce an exponential variation in entrainment and deposition probabilities, while \citet{Sawai1987} concluded that the variation is linear.
For simplicity, we incorporate the scaling of \citet{Sawai1987} and note its equivalence to the \citet{Wong2007} scaling when bed elevation changes are small.
Because experimental distributions of bed elevations are often symmetrical, \citep{Crickmore1962, Pender2001,Wong2007, Martin2014}, we expect the erosion and deposition feedbacks to have the same strength.
That is, as bed elevation changes drive up (down) erosion rates, so they drive down (up) deposition rates to the same degree.
Merging these ideas with those of \citet{Ancey2008}, we write the four possible transitions with local bed elevation-dependent entrainment and deposition rates as
\begin{align}
	&R_{MI}(n+1|n) = \nu && &\text{migration in}, \label{eq:rate1}\\
	&R_E(n+1,m-1|n,m)=(\lambda + \mu n)[1 + \kappa m], && &\text{entrainment}, \label{eq:rate2}\\
	&R_D(n-1,m+1|n,m)=\sigma n [1- \kappa m ], && &\text{deposition}. \label{eq:rate3} \\
	&R_{MO}(n-1|n) =\gamma n && &\text{migration out} \label{eq:rate4}.
\end{align}
In equations (\ref{eq:rate2}) and (\ref{eq:rate3}), $\kappa$ is a coupling constant between bed elevations and the entrainment and deposition rates.
$\nu$ is the rate of migration into the control volume, $\lambda$ is the conventional entrainment rate, $\mu$ is the collective entrainment rate, $\sigma$ is the deposition rate, and $\gamma$ is the rate of migration out of the control volume.
At $m=0$, these equations reduce to those of the \citet{Ancey2008} model.
Away from this elevation, entrainment and deposition are alternatively suppressed and enhanced depending on the sign of $m$, constituting a feedback between bed elevation changes and erosion and deposition.
We refer to $\kappa$ as a coupling constant since it controls the strength of this feedback. We later demonstrate the relationship
\begin{equation}\kappa \approx \big(\frac{z_1}{2l}\big)^2 \label{eq:active}
\end{equation}
where $l$ is a characteristic length scale of bed elevation change that we interpret as the active layer depth \citep{Wong2007,Church2017}.
All four rates are independent of the past history of the populations and depend only on the current populations $(n,m)$. 
As a result, the model is Markovian \citep{Cox1965, VanKampen1992}, meaning time intervals between any two subsequent transitions are exponentially distributed \citep{Gillespie2007}.

>>>>

Resting time distributions have been carefully studied in relation to these models because the predicted diffusion characteristics are critically dependent on whether the distribution has a light or heavy tail \citep{Bradley2017,Martin2012,Weeks1998}.
Resting times have puzzled researchers because early experiments show exponential distributions \citep{Einstein1937,Hubbell1964, Yano1969,Nakagawa1976}, while later experiments show heavy-tailed power-law distributions \citep{Martin2012,Voepel2013, Olinde2015, Pretzlav2016a, Bradley2017, Liu2019}.
A predominant hypothesis is that power-law distributed resting times originate from buried grains \citep{Martin2014,Voepel2013}; this hypothesis permits surface grains to retain exponential resting times. 
Conceptually, when grains rest on the surface, material transported from upstream can deposit on top of them, preventing entrainment until its removal, driving up resting times and imparting a heavy tail to the distribution.
To our knowledge, \citet{Martin2014} have provided the only direct support for this hypothesis by tracking grains through complete cycles of burial and exhumation using a narrow flume with glass walls.
They observed heavy-tailed resting times of buried grains and described their results with a mathematical model similar to an earlier effort by \citet{Voepel2013}.
Both of these models treat bed elevation changes as a random walk and interpret resting times as return periods from above in the bed elevation time-series \citep{Redner2007}. Each describes resting time distributions from different experiments, but they rely on different random walk models, and their treatment of bed elevations as a process independent of sediment transport is questionable at first glance, since bedload transport is the source of bed elevation changes \citep{Wong2007}, and neither model explicitly includes bedload transport.
Models of sedimentary bed evolution incorporating sediment transport processes might enhance understanding of sediment resting times.

The second type of stochastic model prescribes rates (probabilities per unit time) to the erosion and deposition events of individual grains within a control volume to calculate the particle activity \citep{Einstein1950}.
These approaches aim at a complete statistical characterization of the bedload flux \citep{Furbish2012a,Fathel2015,Furbish2017,Heyman2016}, including probability distributions \citep{Ancey2006,Ancey2008}, spatial and temporal characteristics of its fluctuations \citep{Ancey2008, Heyman2014a, Roseberry2012,Dhont2018}, and the dependence of these statistical characteristics on the length and time scales over which they are measured or calculated \citep{Singh2009,Singh2012,Saletti2015,Ma2014}.
A recent surge in research activity has generated rapid progress and spawned many new inquiries in this subject. 
For example, \citet{Ancey2006} demonstrated that a constant erosion rate as originally proposed by \citet{Einstein1950} was insufficient to develop realistically large particle activity fluctuations, so they added a positive feedback between the particle activity and erosion rate they called ``collective entrainment'' \citep{Ancey2008, Heyman2013,Heyman2014,Ma2014, Lee2018}.
While they deemed this feedback necessary to model realistic activity fluctuations, the implications of this collective entrainment term on bed topography and particle activity changes has not been fully explored.

@@

Bed elevation changes modify the likelihood of entrainment and deposition in a negative feedback \citep{Sawai1987, Wong2007}; that is, aggradation increases the likelihood of entrainment, while degradation increases the likelihood of deposition.
\citet{Wong2007} concluded that bed elevation changes induce an exponential variation in entrainment and deposition probabilities, while \citet{Sawai1987} concluded that the variation is linear.
For simplicity, we incorporate the scaling of \citet{Sawai1987} and note its equivalence to the \citet{Wong2007} scaling when bed elevation changes are small.
Because experimental distributions of bed elevations are often symmetrical, \citep{Crickmore1962, Pender2001,Wong2007, Martin2014}, we expect the erosion and deposition feedbacks to have the same strength.
That is, as bed elevation changes drive up (down) erosion rates, so they drive down (up) deposition rates to the same degree.
Merging these ideas with those of \citet{Ancey2008}, we write the four possible transitions with local bed elevation-dependent entrainment and deposition rates as
\begin{align}
	&R_{MI}(n+1|n) = \nu && &\text{migration in}, \label{eq:rate1}\\
	&R_E(n+1,m-1|n,m)=(\lambda + \mu n)[1 + \kappa m], && &\text{entrainment}, \label{eq:rate2}\\
	&R_D(n-1,m+1|n,m)=\sigma n [1- \kappa m ], && &\text{deposition}. \label{eq:rate3} \\
	&R_{MO}(n-1|n) =\gamma n && &\text{migration out} \label{eq:rate4}.
\end{align}
In equations (\ref{eq:rate2}) and (\ref{eq:rate3}), $\kappa$ is a coupling constant between bed elevations and the entrainment and deposition rates.
$\nu$ is the rate of migration into the control volume, $\lambda$ is the conventional entrainment rate, $\mu$ is the collective entrainment rate, $\sigma$ is the deposition rate, and $\gamma$ is the rate of migration out of the control volume.
At $m=0$, these equations reduce to those of the \citet{Ancey2008} model.
Away from this elevation, entrainment and deposition are alternatively suppressed and enhanced depending on the sign of $m$, constituting a feedback between bed elevation changes and erosion and deposition.
We refer to $\kappa$ as a coupling constant since it controls the strength of this feedback. We later demonstrate the relationship
\begin{equation}\kappa \approx \big(\frac{z_1}{2l}\big)^2 \label{eq:active}
\end{equation}
where $l$ is a characteristic length scale of bed elevation change that we interpret as the active layer depth \citep{Wong2007,Church2017}.
All four rates are independent of the past history of the populations and depend only on the current populations $(n,m)$. 
As a result, the model is Markovian \citep{Cox1965, VanKampen2007}, meaning time intervals between any two subsequent transitions are exponentially distributed \citep{Gillespie2007}.
